% 英文摘要

Marine propellers are susceptible to complex vibration responses during operation due to non-uniform wake, blade-frequency excitation, and fluid–structure interaction. Their modal characteristics are directly related to the structural safety, vibration noise level, and service reliability of the propulsion system. Traditional finite element methods can calculate propeller modal frequencies with relatively high accuracy. However, in multi-parameter design and large-scale scheme evaluation, repeated geometric modeling, mesh generation, and finite element solution are required, resulting in high computational cost and low efficiency. Therefore, it is difficult to meet the demand for rapid prediction. To address this problem, this thesis proposes a rapid prediction method for the wet modal frequencies of marine propellers by combining high-precision finite element simulation, modal test validation, and machine learning surrogate models.

In terms of high-precision finite element simulation, based on propeller modal analysis theory and finite element theory, a high-precision simulation procedure for dry and wet modal analysis of marine propellers is established. The key procedures, including element type selection, mesh size determination, material parameter setting and correction, and structural–fluid coupled modeling for wet modal analysis, are clarified. Taking the DTMB4382 propeller as an example, the first 20 dry and wet modal frequencies are calculated, and the corresponding added-water frequency correction coefficients are obtained, providing a basis for subsequent rapid prediction of wet modal frequencies. In terms of database construction, parameterized modeling and automated simulation techniques are combined. Python and CadQuery are used to generate three-dimensional propeller models, HyperMesh batch scripts are used for automatic mesh generation, and Abaqus Python scripts are used for modal calculation and frequency extraction. On this basis, a “propeller geometric parameters–dry modal frequencies” database is constructed, providing data support for machine learning model training. In terms of modal testing, a dry/wet modal testing procedure based on the moving hammer method, acceleration response frequency response function testing, and modal parameter identification is established. Dry and wet modal tests are carried out in air and water for a seven-bladed propeller and a ten-bladed propeller, respectively. The comparison between the simulation results and test results of the seven-bladed propeller verifies the accuracy and engineering applicability of the proposed simulation procedure.

In terms of machine learning prediction, a dry modal frequency prediction model is developed based on the established “propeller geometric parameters–dry modal frequencies” database. First, Pearson correlation coefficients and Spearman correlation coefficients are used to analyze the correlation between the original geometric parameters and the first 20 dry modal frequencies. A total of 24 key geometric parameters, including propeller diameter, blade number, and the chord length and thickness at 11 radial sections, are selected as model inputs, thereby realizing input feature dimensionality reduction. Then, four machine learning models are constructed to predict the first 20 dry modal frequencies of the propeller, including Random Forest, Gradient Boosting Regression, Gaussian Process Regression, and Multi-Layer Perceptron. Random search is used for hyperparameter optimization. The prediction performance of the models is evaluated using four indicators: MAE, RMSE, R², and MAPE. The results show that the Multi-Layer Perceptron model achieves the best overall performance, with an MAE of 5.133937, an RMSE of 7.445019, an R² of 0.996914, and a MAPE of 1.889931\% on the test set. This indicates that the model can accurately learn the nonlinear mapping relationship between key propeller geometric parameters and dry modal frequencies. On this basis, the first 20 dry modal frequencies predicted by the Multi-Layer Perceptron model are combined with the added-water frequency correction coefficients, forming a rapid wet modal frequency prediction process of “geometric parameter input–dry modal frequency prediction–added-water coefficient correction–wet modal frequency output.”